---
title: "Mapping global high-resolution sustainable development trends over two decades"
author: "Yangyang Guo, Wenhua Yu, Rongbin Xu, Zhihu Xu, Tingting Ye, Michael Tong, Sotiris Vardoulakis, Yuming Guo, and Xuemei Bai"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

required_pkgs <- c(
  "tidyverse", "dplyr", "broom", "raster", "stars", "rasterVis",
  "sf", "sp", "rnaturalearth", "furrr", "future",
  "cowplot", "openxlsx", "Kendall"
)

missing_pkgs <- required_pkgs[!vapply(required_pkgs, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing_pkgs) > 0) {
  stop(
    "Please install the required packages before running this script: ",
    paste(missing_pkgs, collapse = ", "),
    call. = FALSE
  )
}

invisible(lapply(required_pkgs, library, character.only = TRUE))

root_dir <- "."
dir_data <- file.path(root_dir, "data")
dir_grid <- file.path(dir_data, "grid_data_2000_2019")
dir_external <- file.path(dir_data, "external")
dir_corr <- file.path(dir_data, "correlation")

dir_plots <- file.path(root_dir, "plots_data")
dir_results <- file.path(root_dir, "Results")
dir_results_data <- file.path(dir_results, "data")
dir_results_plots <- file.path(dir_results, "result_plots")
dir_results_corr <- file.path(dir_results, "raster_correlation")

dir_fig3 <- file.path(dir_plots, "Fig_3")
dir_table1 <- file.path(dir_plots, "Table1")
dir_fig4 <- file.path(dir_plots, "Fig4")
dir_ssp_population <- file.path(dir_external, "ssp_population")

dirs_to_create <- c(
  dir_data, dir_grid, dir_external, dir_corr,
  dir_plots, dir_results, dir_results_data,
  dir_results_plots, dir_results_corr,
  dir_fig3, dir_table1, dir_fig4, dir_ssp_population
)

invisible(lapply(dirs_to_create, dir.create, recursive = TRUE, showWarnings = FALSE))

check_file <- function(path) {
  if (!file.exists(path)) {
    stop("Required file not found: ", normalizePath(path, winslash = "/", mustWork = FALSE), call. = FALSE)
  }
  path
}

path_grid <- function(indicator) {
  file.path(dir_grid, paste0("grid_", indicator, "_data_df_update.rds"))
}

path_result_trend <- function(indicator) {
  file.path(dir_results, paste0(indicator, "_trend_global_update.tif"))
}
```

This repository contains the main code for the article **"Mapping global high-resolution sustainable development trends over two decades"**.

We use four headline indicators to measure sustainable development trends over the past two decades:

- **PM2.5** – an air pollution control indicator
- **CO2** – a climate change mitigation indicator
- **Nightlight** – a socioeconomic activity and infrastructure development indicator
- **NDVI** – an ecosystem vitality indicator

The annual indicator data used in this analysis are available via Figshare. In this public code version, all file paths use relative project directories.

## Expected project structure

```{r project-structure, echo=FALSE, results='asis'}
cat("
repo/
├─ code.Rmd
├─ data/
│  ├─ grid_data_2000_2019/
│  │  ├─ grid_PM2_5_data_df_update.rds
│  │  ├─ grid_co2_data_df_update.rds
│  │  ├─ grid_NDVI_data_df_update.rds
│  │  └─ grid_nightlight_data_df_update.rds
│  ├─ correlation/
│  │  ├─ Sustainable_trend_pattern_with_GDP_extracted.rds
│  │  └─ table1_df_updated_20250307.rds
│  └─ external/
│     └─ ssp_population/
│        ├─ ssp1_total_2030.nc4
│        ├─ ssp2_total_2030.nc4
│        ├─ ssp3_total_2030.nc4
│        ├─ ssp4_total_2030.nc4
│        └─ ssp5_total_2030.nc4
├─ plots_data/
│  ├─ landscan_global_UN_adjusted_2019.tif
│  ├─ year_urban_2019_resample.tif
│  ├─ World_bank_income_group_2019.csv
│  └─ region_raster_df.rds
└─ Results/
")
```

## 1. Change in indicator values per decade

```{r}
trend_function_update <- function(indicator = "co2") {

  grid_file <- path_grid(indicator)
  grid_data_df <- readRDS(check_file(grid_file))

  e1 <- Sys.time()

  trend_data <-
    grid_data_df %>%
    tidyr::nest(data = -c(x, y)) %>%
    dplyr::mutate(
      model = purrr::map(data, ~ stats::lm(value ~ year, data = .x)),
      tidied = purrr::map(model, broom::tidy)
    ) %>%
    tidyr::unnest(tidied) %>%
    dplyr::filter(term == "year") %>%
    dplyr::select(x, y, estimate, p.value)

  saveRDS(
    trend_data,
    file = file.path(dir_results_data, paste0("trend_", indicator, "_update2.rds"))
  )

  e2 <- Sys.time()
  print(e2 - e1)

  trend_data_clean <- trend_data %>%
    dplyr::filter(
      estimate >= stats::quantile(trend_data$estimate, 0.0001, na.rm = TRUE),
      estimate <= stats::quantile(trend_data$estimate, 0.9999, na.rm = TRUE)
    ) %>%
    dplyr::mutate(estimate = estimate * 10) %>%
    dplyr::select(x, y, estimate)

  raster_trend_data_clean <-
    raster::rasterFromXYZ(
      xyz = trend_data_clean,
      crs = "+proj=longlat +datum=WGS84 +no_defs"
    )

  raster::writeRaster(
    x = raster_trend_data_clean,
    filename = path_result_trend(indicator),
    format = "GTiff",
    overwrite = TRUE
  )
}
```

```{r}
trend_function_update(indicator = "PM2_5")
trend_function_update(indicator = "co2")
trend_function_update(indicator = "NDVI")
trend_function_update(indicator = "nightlight")
```

## 2. Combine indicators

```{r}
PM2_5_trend <- raster::raster(check_file(path_result_trend("PM2_5")))
co2_trend <- raster::raster(check_file(path_result_trend("co2")))
nightlight_trend <- raster::raster(check_file(path_result_trend("nightlight")))
NDVI_trend <- raster::raster(check_file(path_result_trend("NDVI")))

new_extent <- raster::extent(-180, 180, -55.4, 74)

co2_trend_crop <- raster::crop(co2_trend, new_extent)
nightlight_trend_crop <- raster::crop(nightlight_trend, new_extent)
NDVI_trend_crop <- raster::crop(NDVI_trend, new_extent)

raster_stack_co2_nl_ndvi <- raster::stack(co2_trend_crop, nightlight_trend_crop, NDVI_trend_crop)
names(raster_stack_co2_nl_ndvi) <- c("co2_trend", "nightlight_trend", "NDVI_trend")

PM2_5_trend_crop <- raster::resample(PM2_5_trend, raster_stack_co2_nl_ndvi, method = "bilinear")

raster_stack_all <- raster::stack(raster_stack_co2_nl_ndvi, PM2_5_trend_crop)
names(raster_stack_all) <- c("co2_trend", "nightlight_trend", "NDVI_trend", "PM2_5_trend")

rm(co2_trend_crop, nightlight_trend_crop, NDVI_trend_crop, PM2_5_trend_crop)
rm(NDVI_trend, nightlight_trend, co2_trend, PM2_5_trend, raster_stack_co2_nl_ndvi)

raster_stack_all_stars <- stars::st_as_stars(raster_stack_all)
stars::write_stars(raster_stack_all_stars, file.path(dir_results, "raster_stack_all_update.tif"))
```

## 3. Classification of trend

```{r}
raster_stack_all <- raster::stack(check_file(file.path(dir_results, "raster_stack_all_update.tif")))

raster_stack_all <- raster::rasterToPoints(raster_stack_all)
raster_stack_all_df <- as.data.frame(raster_stack_all, xy = TRUE) %>%
  tidyr::drop_na()

raster_stack_all_df$co2_score <- ifelse(raster_stack_all_df$co2_trend <= 0, 1, 0)
raster_stack_all_df$PM2_5_score <- ifelse(raster_stack_all_df$PM2_5_trend <= 0, 1, 0)
raster_stack_all_df$nightlight_score <- ifelse(raster_stack_all_df$nightlight_trend > 0, 1, 0)
raster_stack_all_df$NDVI_score <- ifelse(raster_stack_all_df$NDVI_trend > 0, 1, 0)

raster_stack_all_df$score <-
  raster_stack_all_df$co2_score +
  raster_stack_all_df$nightlight_score +
  raster_stack_all_df$NDVI_score +
  raster_stack_all_df$PM2_5_score

raster_stack_all_df$score <- as.factor(raster_stack_all_df$score)

saveRDS(raster_stack_all_df, file = file.path(dir_results, "raster_stack_all_df_update.rds"))

raster_score <- raster_stack_all_df %>%
  dplyr::select(x, y, score) %>%
  dplyr::mutate(score = as.numeric(as.character(score))) %>%
  raster::rasterFromXYZ(xyz = ., crs = "+proj=longlat +datum=WGS84 +no_defs")

raster::writeRaster(
  x = raster_score,
  filename = file.path(dir_results, "raster_score_update.tif"),
  format = "GTiff",
  overwrite = TRUE
)
```

## 4. Plot raster score

```{r}
p_score <- ggplot2::ggplot(data = raster_stack_all_df, ggplot2::aes(x = x, y = y, fill = score)) +
  ggplot2::geom_raster() +
  ggplot2::coord_fixed() +
  ggplot2::scale_fill_manual(
    values = c("#e12729", "#f37324", "#f8cc1b", "#72b043", "#007f4e"),
    name = "Classification",
    labels = c("Very Bad", "Bad", "Medium", "Good", "Very Good")
  ) +
  ggplot2::theme_minimal() +
  ggplot2::labs(fill = "score", x = NULL, y = NULL)

tiff(file.path(dir_results_plots, "plot_score_update.tif"), width = 2500, height = 1000, units = "px", res = 300)
print(p_score)
dev.off()
```

```{r}
p_co2 <- ggplot2::ggplot(data = raster_stack_all_df, ggplot2::aes(x = x, y = y, fill = as.factor(co2_score))) +
  ggplot2::geom_raster() +
  ggplot2::coord_fixed() +
  ggplot2::scale_fill_manual(
    values = c("#e12729", "#72b043"),
    name = "Trend",
    labels = c("Increase", "Decline")
  ) +
  ggplot2::theme_minimal() +
  ggplot2::labs(fill = "co2_score", x = NULL, y = NULL) +
  ggplot2::ggtitle("CO2")

p_PM25 <- ggplot2::ggplot(data = raster_stack_all_df, ggplot2::aes(x = x, y = y, fill = as.factor(PM2_5_score))) +
  ggplot2::geom_raster() +
  ggplot2::coord_fixed() +
  ggplot2::scale_fill_manual(
    values = c("#e12729", "#72b043"),
    name = "Trend",
    labels = c("Increase", "Decline")
  ) +
  ggplot2::theme_minimal() +
  ggplot2::labs(fill = "PM2_5_score", x = NULL, y = NULL) +
  ggplot2::ggtitle("PM2.5")

p_nightlight <- ggplot2::ggplot(data = raster_stack_all_df, ggplot2::aes(x = x, y = y, fill = as.factor(nightlight_score))) +
  ggplot2::geom_raster() +
  ggplot2::coord_fixed() +
  ggplot2::scale_fill_manual(
    values = c("#e12729", "#72b043"),
    name = "Trend",
    labels = c("Decline", "Increase")
  ) +
  ggplot2::theme_minimal() +
  ggplot2::labs(fill = "nightlight_score", x = NULL, y = NULL) +
  ggplot2::ggtitle("Nightlight")

p_NDVI <- ggplot2::ggplot(data = raster_stack_all_df, ggplot2::aes(x = x, y = y, fill = as.factor(NDVI_score))) +
  ggplot2::geom_raster() +
  ggplot2::coord_fixed() +
  ggplot2::scale_fill_manual(
    values = c("#e12729", "#72b043"),
    name = "Trend",
    labels = c("Decline", "Increase")
  ) +
  ggplot2::theme_minimal() +
  ggplot2::labs(fill = "NDVI_score", x = NULL, y = NULL) +
  ggplot2::ggtitle("NDVI")
```

```{r}
p_all <- cowplot::plot_grid(p_co2, p_PM25, p_nightlight, p_NDVI, ncol = 2, nrow = 2)

tiff(file.path(dir_results_plots, "plot_4indicators_update.tif"), width = 5000, height = 2000, units = "px", res = 300)
print(p_all)
dev.off()
```

## 5. Plot trend of each indicator with exact value

```{r}
p_co2_value <- ggplot2::ggplot(data = raster_stack_all_df, ggplot2::aes(x = x, y = y, fill = co2_trend)) +
  ggplot2::geom_raster() +
  ggplot2::coord_fixed() +
  ggplot2::scale_fill_viridis_c(option = "D") +
  ggplot2::theme_minimal() +
  ggplot2::labs(fill = "CO2 trend", x = NULL, y = NULL) +
  ggplot2::ggtitle("CO2")

p_PM25_value <- ggplot2::ggplot(data = raster_stack_all_df, ggplot2::aes(x = x, y = y, fill = PM2_5_trend)) +
  ggplot2::geom_raster() +
  ggplot2::coord_fixed() +
  ggplot2::scale_fill_viridis_c(option = "D") +
  ggplot2::theme_minimal() +
  ggplot2::labs(fill = "PM2.5 trend", x = NULL, y = NULL) +
  ggplot2::ggtitle("PM2.5")

p_nightlight_value <- ggplot2::ggplot(data = raster_stack_all_df, ggplot2::aes(x = x, y = y, fill = nightlight_trend)) +
  ggplot2::geom_raster() +
  ggplot2::coord_fixed() +
  ggplot2::scale_fill_viridis_c(option = "D", direction = -1) +
  ggplot2::theme_minimal() +
  ggplot2::labs(fill = "Nightlight trend", x = NULL, y = NULL) +
  ggplot2::ggtitle("Nightlight")

p_NDVI_value <- ggplot2::ggplot(data = raster_stack_all_df, ggplot2::aes(x = x, y = y, fill = NDVI_trend)) +
  ggplot2::geom_raster() +
  ggplot2::coord_fixed() +
  ggplot2::scale_fill_viridis_c(option = "D", direction = -1) +
  ggplot2::theme_minimal() +
  ggplot2::labs(fill = "NDVI trend", x = NULL, y = NULL) +
  ggplot2::ggtitle("NDVI")
```

```{r}
p_all_value <- cowplot::plot_grid(p_co2_value, p_PM25_value, p_nightlight_value, p_NDVI_value, ncol = 2, nrow = 2)

tiff(file.path(dir_results_plots, "plot_4indicators_value_update.tif"), width = 5000, height = 2000, units = "px", res = 300)
print(p_all_value)
dev.off()
```

## 6. Category with 16 groups

1. Population in 2019  
2. Population in 2030 (SSP1–SSP5)  
3. Category with 16 groups  
4. Population totals and proportions  

```{r}
raster_stack_all_df <- readRDS(check_file(file.path(dir_results, "raster_stack_all_df_update.rds")))

raster_stack_all_df <- raster_stack_all_df %>%
  dplyr::mutate(
    combined_category = paste0(co2_score, PM2_5_score, nightlight_score, NDVI_score),
    combined_category = factor(
      combined_category,
      levels = c("0000", "1000", "0100", "0010", "0001", "1100", "1010", "1001",
                 "0110", "0101", "0011", "1110", "1101", "1011", "0111", "1111")
    ),
    combined_type = as.integer(combined_category)
  )

raster_category <- raster_stack_all_df %>%
  dplyr::select(x, y, combined_type) %>%
  raster::rasterFromXYZ(xyz = ., crs = "+proj=longlat +datum=WGS84 +no_defs")

raster::writeRaster(
  x = raster_category,
  filename = file.path(dir_plots, "raster_category_16_update.tif"),
  format = "GTiff",
  overwrite = TRUE
)
```

```{r}
year_data_df_2019 <- raster::stack(check_file(file.path(dir_plots, "landscan_global_UN_adjusted_2019.tif")))
r <- year_data_df_2019

aggregation_factor <- round(0.1 / raster::res(r)[1])

r_aggregated <- raster::aggregate(r, fact = aggregation_factor, fun = sum, na.rm = TRUE)

year_pop_2019_resample <- raster::resample(r_aggregated, raster_category, method = "ngb")

print(raster::cellStats(year_pop_2019_resample, stat = sum))

raster::writeRaster(
  x = year_pop_2019_resample,
  filename = file.path(dir_plots, "year_pop_2019_resample_UN_adjusted.tif"),
  format = "GTiff",
  overwrite = TRUE
)

for (i in 1:5) {
  ssp_file <- file.path(dir_ssp_population, paste0("ssp", i, "_total_2030.nc4"))
  year_pop_ssp_2030 <- raster::raster(check_file(ssp_file))

  aggregation_factor_ssp <- round(0.1 / raster::res(year_pop_ssp_2030)[1])

  year_pop_ssp_2030_0.1 <- raster::aggregate(
    year_pop_ssp_2030,
    fact = aggregation_factor_ssp,
    fun = sum,
    na.rm = TRUE
  )

  year_pop_ssp_2030_resample <- raster::resample(year_pop_ssp_2030_0.1, raster_category, method = "bilinear")

  raster::writeRaster(
    x = year_pop_ssp_2030_resample,
    filename = file.path(dir_plots, paste0("year_pop_ssp", i, "_2030_resample.tif")),
    format = "GTiff",
    overwrite = TRUE
  )
}
```

```{r}
pop_2019 <- raster::raster(check_file(file.path(dir_plots, "year_pop_2019_resample_UN_adjusted.tif")))

pop_2030_ssp1 <- raster::raster(check_file(file.path(dir_plots, "year_pop_ssp1_2030_resample.tif")))
pop_2030_ssp2 <- raster::raster(check_file(file.path(dir_plots, "year_pop_ssp2_2030_resample.tif")))
pop_2030_ssp3 <- raster::raster(check_file(file.path(dir_plots, "year_pop_ssp3_2030_resample.tif")))
pop_2030_ssp4 <- raster::raster(check_file(file.path(dir_plots, "year_pop_ssp4_2030_resample.tif")))
pop_2030_ssp5 <- raster::raster(check_file(file.path(dir_plots, "year_pop_ssp5_2030_resample.tif")))

category_type <- raster::raster(check_file(file.path(dir_plots, "raster_category_16_update.tif")))

raster_category_pop_2019_2030 <-
  raster::stack(c(category_type, pop_2019, pop_2030_ssp1, pop_2030_ssp2, pop_2030_ssp3, pop_2030_ssp4, pop_2030_ssp5))

names(raster_category_pop_2019_2030) <- c(
  "category", "pop_2019", "pop_2030_ssp1", "pop_2030_ssp2",
  "pop_2030_ssp3", "pop_2030_ssp4", "pop_2030_ssp5"
)

Fig3_df <- raster::as.data.frame(raster_category_pop_2019_2030, xy = TRUE) %>%
  dplyr::filter(!is.na(category))

saveRDS(Fig3_df, file.path(dir_plots, "Fig_df_update.rds"))

Fig3 <- Fig3_df %>%
  dplyr::mutate(category = as.factor(category)) %>%
  dplyr::group_by(category) %>%
  dplyr::summarise(
    sum_pop_2019 = sum(pop_2019, na.rm = TRUE),
    sum_pop_2030_ssp1 = sum(pop_2030_ssp1, na.rm = TRUE),
    sum_pop_2030_ssp2 = sum(pop_2030_ssp2, na.rm = TRUE),
    sum_pop_2030_ssp3 = sum(pop_2030_ssp3, na.rm = TRUE),
    sum_pop_2030_ssp4 = sum(pop_2030_ssp4, na.rm = TRUE),
    sum_pop_2030_ssp5 = sum(pop_2030_ssp5, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    proportion_2019 = round((sum_pop_2019 / sum(sum_pop_2019, na.rm = TRUE)) * 100, 2),
    proportion_2030_ssp1 = round((sum_pop_2030_ssp1 / sum(sum_pop_2030_ssp1, na.rm = TRUE)) * 100, 2),
    proportion_2030_ssp2 = round((sum_pop_2030_ssp2 / sum(sum_pop_2030_ssp2, na.rm = TRUE)) * 100, 2),
    proportion_2030_ssp3 = round((sum_pop_2030_ssp3 / sum(sum_pop_2030_ssp3, na.rm = TRUE)) * 100, 2),
    proportion_2030_ssp4 = round((sum_pop_2030_ssp4 / sum(sum_pop_2030_ssp4, na.rm = TRUE)) * 100, 2),
    proportion_2030_ssp5 = round((sum_pop_2030_ssp5 / sum(sum_pop_2030_ssp5, na.rm = TRUE)) * 100, 2)
  )

write.csv(Fig3, file = file.path(dir_fig3, "Fig3_update_ssp1_5_update_UN_adjusted.csv"), row.names = FALSE)
```

## 7. Number and proportion of best and worst categories

Rows include: global, continent, income group, and urban/rural.

```{r}
pop_2019 <- raster::raster(check_file(file.path(dir_plots, "year_pop_2019_resample_UN_adjusted.tif")))
category_type <- raster::raster(check_file(file.path(dir_plots, "raster_category_16_update.tif")))
urban_2019 <- raster::raster(check_file(file.path(dir_plots, "year_urban_2019_resample.tif")))
income_group <- read.csv(check_file(file.path(dir_plots, "World_bank_income_group_2019.csv")))
pop_2030 <- raster::raster(check_file(file.path(dir_plots, "year_pop_ssp2_2030_resample.tif")))

raster_category_pop_urban <-
  raster::stack(c(category_type, pop_2019, pop_2030, urban_2019))

names(raster_category_pop_urban) <- c("category_type", "pop_2019", "pop_2030_ssp2", "urban_code_2019")

raster_category_pop_urban <- raster::as.data.frame(raster_category_pop_urban, xy = TRUE) %>%
  dplyr::mutate(
    x = round(x, 2),
    y = round(y, 2)
  )
```

```{r}
region_raster_df <- readRDS(check_file(file.path(dir_plots, "region_raster_df.rds"))) %>%
  dplyr::mutate(
    X = round(X, 2),
    Y = round(Y, 2)
  )

sub_region_category_intersect <- raster_category_pop_urban %>%
  dplyr::left_join(region_raster_df, by = c("x" = "X", "y" = "Y"))
```

```{r}
table1_df <- sub_region_category_intersect %>%
  dplyr::filter(!is.na(category_type)) %>%
  dplyr::left_join(income_group, by = c("iso_a3" = "ISO3")) %>%
  dplyr::mutate(
    category_type = factor(category_type, levels = 1:16),
    SUBREGION2 = factor(
      SUBREGION2,
      levels = c(
        "Northern America", "Latin America and the Caribbean", "Northern Europe",
        "Eastern Europe", "Western Europe", "Southern Europe", "Northern Africa",
        "Sub-Saharan Africa", "South-eastern Asia", "Western Asia", "Central Asia",
        "Southern Asia", "Eastern Asia", "Australia and New Zealand",
        "Other regions in Oceania"
      )
    ),
    REGION = factor(REGION, levels = c("Americas", "Europe", "Africa", "Asia", "Oceania")),
    Income2019 = factor(Income2019, levels = c("H", "UM", "LM", "L")),
    urban_2019 = dplyr::if_else(
      urban_code_2019 == 30, "Urban centre",
      dplyr::if_else(urban_code_2019 == 10, "Water", "Rural")
    ),
    urban_2019 = factor(urban_2019, levels = c("Urban centre", "Rural", "Water"))
  )

saveRDS(table1_df, file = file.path(dir_plots, "table1_df_updated.rds"))
```

```{r}
df <- table1_df %>% tidyr::drop_na()

summary_table <- df %>%
  dplyr::mutate(Global = "Global") %>%
  tidyr::pivot_longer(
    cols = c("Global", "REGION", "Income2019", "urban_2019"),
    names_to = "Category",
    values_to = "Value"
  ) %>%
  dplyr::group_by(Category, Value, category_type) %>%
  dplyr::summarise(Total_Population = sum(pop_2019, na.rm = TRUE), .groups = "drop") %>%
  dplyr::group_by(Category, Value) %>%
  dplyr::mutate(
    Total_Sum = sum(Total_Population, na.rm = TRUE),
    Proportion = round(Total_Population / Total_Sum * 100, 2)
  ) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(
    names_from = category_type,
    values_from = c(Total_Population, Proportion),
    names_sep = "_"
  )

write.csv(summary_table, file = file.path(dir_table1, "summary_table1_update.csv"), row.names = FALSE)
```

## 8. Rank countries by best and worst categories

A. Top countries by population in best category  
B. Top countries by population in worst category  
C. Top countries by proportion in best category  
D. Top countries by proportion in worst category  

```{r}
table1_df <- readRDS(check_file(file.path(dir_plots, "table1_df_updated.rds")))

Fig4_data <- table1_df %>%
  dplyr::group_by(iso_a3, Country_name, category_type) %>%
  dplyr::summarise(sum_pop = sum(pop_2019, na.rm = TRUE), .groups = "drop_last") %>%
  dplyr::group_by(iso_a3, Country_name) %>%
  dplyr::mutate(
    proportion = sum_pop / sum(sum_pop, na.rm = TRUE),
    proportion = round(proportion * 100, 2)
  ) %>%
  dplyr::ungroup()

saveRDS(Fig4_data, file = file.path(dir_fig4, "Fig4_data_update.rds"))

top15_best_numb <- Fig4_data %>%
  dplyr::filter(category_type == 16) %>%
  dplyr::arrange(dplyr::desc(sum_pop)) %>%
  dplyr::slice(1:15)

top15_worst_numb <- Fig4_data %>%
  dplyr::filter(category_type == 1) %>%
  dplyr::arrange(dplyr::desc(sum_pop)) %>%
  dplyr::slice(1:15)

top15_best_proportion <- Fig4_data %>%
  dplyr::filter(category_type == 16) %>%
  dplyr::arrange(dplyr::desc(proportion)) %>%
  dplyr::slice(1:15)

top15_worst_proportion <- Fig4_data %>%
  dplyr::filter(category_type == 1) %>%
  dplyr::arrange(dplyr::desc(proportion)) %>%
  dplyr::slice(1:15)

write.csv(top15_best_numb, file = file.path(dir_fig4, "top15_best_numb_update.csv"), row.names = FALSE)
write.csv(top15_worst_numb, file = file.path(dir_fig4, "top15_worst_numb_update.csv"), row.names = FALSE)
write.csv(top15_best_proportion, file = file.path(dir_fig4, "top15_best_proportion_update.csv"), row.names = FALSE)
write.csv(top15_worst_proportion, file = file.path(dir_fig4, "top15_worst_proportion_update.csv"), row.names = FALSE)
```

## 9. Grid-cell Kendall correlation across indicators

```{r}
grid_cell_correlation <- function(
  var1,
  var2,
  grid_all_df,
  use_diff = TRUE,
  output_dir = dir_results_plots,
  raster_dir = dir_results_corr
) {

  diff_tag <- if (use_diff) "diff" else "raw"
  pair_label <- paste0(var1, "_vs_", var2, "_", diff_tag)
  crs_wgs84 <- "+proj=longlat +datum=WGS84 +no_defs"

  dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
  dir.create(raster_dir, showWarnings = FALSE, recursive = TRUE)
  dir.create(dir_results_data, showWarnings = FALSE, recursive = TRUE)

  cell_cor <- grid_all_df %>%
    dplyr::select(x, y, year, dplyr::all_of(c(var1, var2))) %>%
    tidyr::drop_na() %>%
    dplyr::arrange(x, y, year) %>%
    tidyr::nest(data = -c(x, y)) %>%
    dplyr::mutate(
      ct = furrr::future_map(
        data,
        function(d) {
          v1 <- if (use_diff) diff(d[[var1]]) else d[[var1]]
          v2 <- if (use_diff) diff(d[[var2]]) else d[[var2]]

          if (length(v1) < 2 || length(v2) < 2) {
            return(list(estimate = NA_real_, p.value = NA_real_))
          }

          test_res <- tryCatch(
            stats::cor.test(v1, v2, method = "kendall"),
            error = function(e) NULL
          )

          if (is.null(test_res)) {
            list(estimate = NA_real_, p.value = NA_real_)
          } else {
            list(
              estimate = unname(test_res$estimate),
              p.value = test_res$p.value
            )
          }
        },
        .options = furrr::furrr_options(seed = TRUE)
      ),
      tau = purrr::map_dbl(ct, ~ .x$estimate),
      p_value = purrr::map_dbl(ct, ~ .x$p.value)
    ) %>%
    dplyr::select(x, y, tau, p_value)

  saveRDS(cell_cor, file.path(dir_results_data, paste0("cell_cor_", pair_label, ".rds")))

  raster::writeRaster(
    raster::rasterFromXYZ(cell_cor[, c("x", "y", "tau")], crs = crs_wgs84),
    file.path(raster_dir, paste0("tau_", pair_label, ".tif")),
    overwrite = TRUE
  )

  raster::writeRaster(
    raster::rasterFromXYZ(cell_cor[, c("x", "y", "p_value")], crs = crs_wgs84),
    file.path(raster_dir, paste0("pval_", pair_label, ".tif")),
    overwrite = TRUE
  )

  p_tau <- ggplot2::ggplot(cell_cor, ggplot2::aes(x, y, fill = tau)) +
    ggplot2::geom_raster() +
    ggplot2::coord_fixed() +
    ggplot2::scale_fill_viridis_c(option = "D", limits = c(-1, 1), name = "tau") +
    ggplot2::theme_minimal() +
    ggplot2::labs(title = paste0(var1, " vs ", var2), x = NULL, y = NULL)

  p_pval <- ggplot2::ggplot(cell_cor, ggplot2::aes(x, y, fill = p_value)) +
    ggplot2::geom_raster() +
    ggplot2::coord_fixed() +
    ggplot2::scale_fill_viridis_c(option = "B", direction = -1, name = "p-value") +
    ggplot2::theme_minimal() +
    ggplot2::labs(title = paste0(var1, " vs ", var2), x = NULL, y = NULL)

  tiff(file.path(output_dir, paste0("cor_tau_", pair_label, ".tif")), 2500, 1000, res = 300)
  print(p_tau)
  dev.off()

  tiff(file.path(output_dir, paste0("cor_pval_", pair_label, ".tif")), 2500, 1000, res = 300)
  print(p_pval)
  dev.off()

  list(tau_map = p_tau, pval_map = p_pval, data = cell_cor)
}

run_grid_correlations <- function(
  indicators = c("PM2_5", "co2", "nightlight", "NDVI"),
  output_dir = dir_results_plots
) {
  dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

  n_workers <- max(1L, future::availableCores() - 1L)
  future::plan(future::multisession, workers = n_workers)

  grid_list <- lapply(indicators, function(ind) {
    df <- readRDS(check_file(path_grid(ind)))
    df <- dplyr::mutate(df, x = round(x, 2), y = round(y, 2))
    names(df)[names(df) == "value"] <- ind
    df
  })

  grid_all_df <- Reduce(function(a, b) dplyr::inner_join(a, b, by = c("x", "y", "year")), grid_list)
  pairs <- combn(indicators, 2, simplify = FALSE)

  run_one <- function(use_diff) {
    tag <- if (use_diff) "diff" else "raw"

    res <- lapply(pairs, function(p) {
      grid_cell_correlation(p[1], p[2], grid_all_df, use_diff, output_dir)
    })

    p_tau_all <- cowplot::plot_grid(plotlist = lapply(res, `[[`, "tau_map"), ncol = 2)
    tiff(file.path(output_dir, paste0("cor_tau_all_pairs_", tag, ".tif")), 5000, 3000, res = 300)
    print(p_tau_all)
    dev.off()

    p_pval_all <- cowplot::plot_grid(plotlist = lapply(res, `[[`, "pval_map"), ncol = 2)
    tiff(file.path(output_dir, paste0("cor_pval_all_pairs_", tag, ".tif")), 5000, 3000, res = 300)
    print(p_pval_all)
    dev.off()

    res
  }

  list(
    diff = run_one(TRUE),
    raw = run_one(FALSE)
  )
}

cor_results <- run_grid_correlations()
```

## 10. Regional correlation analysis

```{r}
region_lookup <- readRDS(check_file(file.path(dir_plots, "region_raster_df.rds"))) %>%
  dplyr::mutate(
    X = round(X, 2),
    Y = round(Y, 2),
    SUBREGION2 = ifelse(
      SUBREGION2 %in% c("Australia and New Zealand", "Other regions in Oceania"),
      "Oceania",
      SUBREGION2
    )
  )

pop_raster_df <- raster::raster(check_file(file.path(dir_plots, "year_pop_2019_resample_UN_adjusted.tif"))) %>%
  raster::as.data.frame(xy = TRUE) %>%
  dplyr::rename(pop = 3) %>%
  dplyr::mutate(x = round(x, 2), y = round(y, 2))

indicators <- c("PM2_5", "co2", "nightlight", "NDVI")

regional_means_list <- lapply(indicators, function(ind) {
  df <- readRDS(check_file(path_grid(ind))) %>%
    dplyr::mutate(x = round(x, 2), y = round(y, 2)) %>%
    dplyr::left_join(region_lookup, by = c("x" = "X", "y" = "Y")) %>%
    dplyr::left_join(pop_raster_df, by = c("x", "y"))

  simple <- df %>%
    dplyr::group_by(SUBREGION2, year) %>%
    dplyr::summarise(!!ind := mean(value, na.rm = TRUE), .groups = "drop")

  weighted <- df %>%
    dplyr::filter(pop > 0) %>%
    dplyr::group_by(SUBREGION2, year) %>%
    dplyr::summarise(!!ind := stats::weighted.mean(value, pop, na.rm = TRUE), .groups = "drop")

  list(simple = simple, weighted = weighted)
})

names(regional_means_list) <- indicators

merge_ts <- function(type) {
  Reduce(
    function(a, b) dplyr::inner_join(a, b, by = c("SUBREGION2", "year")),
    lapply(indicators, function(i) regional_means_list[[i]][[type]])
  )
}

regional_ts_simple <- merge_ts("simple")
regional_ts_weighted <- merge_ts("weighted")

regional_cor_table <- function(regional_ts, indicators, use_diff) {
  pairs <- combn(indicators, 2, simplify = FALSE)

  res <- lapply(pairs, function(p) {
    var1 <- p[1]
    var2 <- p[2]

    regional_ts %>%
      dplyr::arrange(SUBREGION2, year) %>%
      tidyr::nest(data = -SUBREGION2) %>%
      dplyr::mutate(
        ct = purrr::map(data, function(d) {
          v1 <- if (use_diff) diff(d[[var1]]) else d[[var1]]
          v2 <- if (use_diff) diff(d[[var2]]) else d[[var2]]

          if (length(v1) < 2) return(list(estimate = NA, p.value = NA))

          test <- tryCatch(stats::cor.test(v1, v2, method = "kendall"), error = function(e) NULL)

          if (is.null(test)) list(estimate = NA, p.value = NA)
          else list(estimate = unname(test$estimate), p.value = test$p.value)
        }),
        tau = purrr::map_dbl(ct, ~ .x$estimate),
        p_value = purrr::map_dbl(ct, ~ .x$p.value)
      ) %>%
      dplyr::select(SUBREGION2, tau, p_value) %>%
      dplyr::rename(
        !!paste0(var1, "_vs_", var2, "_tau") := tau,
        !!paste0(var1, "_vs_", var2, "_p_value") := p_value
      )
  })

  Reduce(function(a, b) dplyr::left_join(a, b, by = "SUBREGION2"), res)
}

result_simple_raw <- regional_cor_table(regional_ts_simple, indicators, FALSE)
result_simple_diff <- regional_cor_table(regional_ts_simple, indicators, TRUE)
result_weighted_raw <- regional_cor_table(regional_ts_weighted, indicators, FALSE)
result_weighted_diff <- regional_cor_table(regional_ts_weighted, indicators, TRUE)

write.csv(result_simple_raw, file.path(dir_results_data, "regional_cor_simple_raw.csv"), row.names = FALSE)
write.csv(result_simple_diff, file.path(dir_results_data, "regional_cor_simple_diff.csv"), row.names = FALSE)
write.csv(result_weighted_raw, file.path(dir_results_data, "regional_cor_weighted_raw.csv"), row.names = FALSE)
write.csv(result_weighted_diff, file.path(dir_results_data, "regional_cor_weighted_diff.csv"), row.names = FALSE)

print(result_simple_raw)
print(result_weighted_diff)
```

## 11. Kendall's tau between sustainability trend pattern and GDP metrics

```{r}
input_rds <- check_file(file.path(dir_corr, "Sustainable_trend_pattern_with_GDP_extracted.rds"))

output_dir <- file.path(dir_corr, "Analysis_3_Kendall_tau_outputs")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

output_grid_rds <- file.path(output_dir, "Gridcell_data_for_mapping_and_statistics.rds")
output_excel <- file.path(output_dir, "Kendall_tau_results_sustainability_vs_GDP.xlsx")
output_csv <- file.path(output_dir, "Kendall_tau_results_sustainability_vs_GDP.csv")
output_rds <- file.path(output_dir, "Kendall_tau_results_sustainability_vs_GDP.rds")

df <- readRDS(input_rds)

cat("Data loaded successfully.\n")
cat("Columns in the dataset:\n")
print(names(df))

required_cols <- c(
  "x",
  "y",
  "name",
  "Pattern_code",
  "GDP_per_capita_2000",
  "GDP_per_capita_2019",
  "GDP_per_capita_mean_2000_2019",
  "GDP_per_capita_trend_2000_2019_pct"
)

missing_cols <- setdiff(required_cols, names(df))
if (length(missing_cols) > 0) {
  stop(
    "The following required columns are missing:\n",
    paste(missing_cols, collapse = ", ")
  )
}

valid_levels <- c("SD", "MD", "MT", "MI", "SI")

df_clean <- df %>%
  dplyr::mutate(
    x = as.numeric(x),
    y = as.numeric(y),
    name = toupper(trimws(as.character(name))),
    Pattern_code = as.integer(Pattern_code),
    GDP_per_capita_2000 = as.numeric(GDP_per_capita_2000),
    GDP_per_capita_2019 = as.numeric(GDP_per_capita_2019),
    GDP_per_capita_mean_2000_2019 = as.numeric(GDP_per_capita_mean_2000_2019),
    GDP_per_capita_trend_2000_2019_pct = as.numeric(GDP_per_capita_trend_2000_2019_pct)
  ) %>%
  dplyr::filter(
    name %in% valid_levels,
    Pattern_code %in% 1:5
  )

cat("\nNumber of rows after cleaning:", nrow(df_clean), "\n")

gridcell_data <- df_clean %>%
  dplyr::select(
    x, y, name, Pattern_code,
    GDP_per_capita_2000,
    GDP_per_capita_2019,
    GDP_per_capita_mean_2000_2019,
    GDP_per_capita_trend_2000_2019_pct
  )

saveRDS(gridcell_data, output_grid_rds)

analysis_vars <- list(
  list(var = "GDP_per_capita_2000", label = "GDP per capita in 2000"),
  list(var = "GDP_per_capita_2019", label = "GDP per capita in 2019"),
  list(var = "GDP_per_capita_mean_2000_2019", label = "Mean GDP per capita, 2000-2019"),
  list(var = "GDP_per_capita_trend_2000_2019_pct", label = "Trend in GDP per capita, 2000-2019 (%/year)")
)

run_kendall_fast <- function(data, xvar, yvar = "Pattern_code", label_text = NULL) {
  sub <- data %>%
    dplyr::filter(!is.na(.data[[xvar]]), !is.na(.data[[yvar]]))

  n_obs <- nrow(sub)

  if (n_obs == 0) {
    return(data.frame(
      Variable = ifelse(is.null(label_text), xvar, label_text),
      Variable_name = xvar,
      N = 0,
      Kendall_tau = NA_real_,
      P_value = NA_real_,
      Statistic_S = NA_real_,
      Method = "Kendall::Kendall",
      stringsAsFactors = FALSE
    ))
  }

  x <- sub[[xvar]]
  y <- sub[[yvar]]

  kt <- Kendall::Kendall(x, y)

  data.frame(
    Variable = ifelse(is.null(label_text), xvar, label_text),
    Variable_name = xvar,
    N = n_obs,
    Kendall_tau = as.numeric(kt$tau),
    P_value = as.numeric(kt$sl),
    Statistic_S = as.numeric(kt$S),
    Method = "Kendall::Kendall",
    stringsAsFactors = FALSE
  )
}

cat("\nRunning 4 Kendall tau analyses...\n")

results_list <- lapply(seq_along(analysis_vars), function(i) {
  item <- analysis_vars[[i]]
  cat("  [", i, "/4] Processing:", item$var, "\n", sep = "")

  run_kendall_fast(
    data = df_clean,
    xvar = item$var,
    yvar = "Pattern_code",
    label_text = item$label
  )
})

results_df <- dplyr::bind_rows(results_list) %>%
  dplyr::mutate(
    Direction = dplyr::case_when(
      is.na(Kendall_tau) ~ NA_character_,
      Kendall_tau > 0 ~ "Positive",
      Kendall_tau < 0 ~ "Negative",
      Kendall_tau == 0 ~ "Zero"
    ),
    Significance = dplyr::case_when(
      is.na(P_value) ~ NA_character_,
      P_value < 0.001 ~ "p < 0.001",
      P_value < 0.01 ~ "p < 0.01",
      P_value < 0.05 ~ "p < 0.05",
      TRUE ~ "Not significant"
    )
  )

cat("\n============================================\n")
cat("Kendall tau results:\n")
print(results_df)
cat("============================================\n")

wb <- openxlsx::createWorkbook()

openxlsx::addWorksheet(wb, "Kendall_results")
openxlsx::writeData(wb, "Kendall_results", results_df)

openxlsx::addWorksheet(wb, "README")
openxlsx::writeData(
  wb, "README",
  data.frame(
    Item = c(
      "Input file",
      "Grid-cell output",
      "Outcome variable",
      "Outcome coding",
      "Predictor variables",
      "Method",
      "Important note"
    ),
    Description = c(
      basename(input_rds),
      basename(output_grid_rds),
      "Sustainable trend pattern",
      "Pattern_code: SD=1, MD=2, MT=3, MI=4, SI=5",
      "GDP_per_capita_2000; GDP_per_capita_2019; GDP_per_capita_mean_2000_2019; GDP_per_capita_trend_2000_2019_pct",
      "Kendall::Kendall",
      "Tau and p-value are global statistics for all valid grid cells, not per-grid-cell values. Per-grid-cell data are saved separately for later mapping."
    ),
    stringsAsFactors = FALSE
  )
)

openxlsx::saveWorkbook(wb, output_excel, overwrite = TRUE)

write.csv(results_df, output_csv, row.names = FALSE, fileEncoding = "UTF-8")
saveRDS(results_df, output_rds)

cat("\nAll Kendall tau analyses completed successfully.\n")
cat("Full grid-cell dataset (.rds) saved to:\n", output_grid_rds, "\n")
cat("Summary Excel saved to:\n", output_excel, "\n")
cat("Summary CSV saved to:\n", output_csv, "\n")
cat("Summary RDS saved to:\n", output_rds, "\n")
```

## 12. Bar plot of Kendall's tau values

```{r}
input_rds <- check_file(file.path(dir_corr, "Analysis_3_Kendall_tau_outputs", "Kendall_tau_results_sustainability_vs_GDP.rds"))

output_dir <- file.path(dir_corr, "Analysis_4_tau_barplot_outputs")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

output_png <- file.path(output_dir, "Barplot_4_Kendall_tau_values_with_pvalues.png")
output_pdf <- file.path(output_dir, "Barplot_4_Kendall_tau_values_with_pvalues.pdf")
output_csv <- file.path(output_dir, "Barplot_4_Kendall_tau_values_with_pvalues_data.csv")
output_rds <- file.path(output_dir, "Barplot_4_Kendall_tau_values_with_pvalues_data.rds")

tau_df <- readRDS(input_rds)

cat("Kendall tau result file loaded successfully.\n")
cat("Columns in the dataset:\n")
print(names(tau_df))

required_cols <- c("Variable_name", "Variable", "Kendall_tau", "P_value", "Direction", "Significance")
missing_cols <- setdiff(required_cols, names(tau_df))

if (length(missing_cols) > 0) {
  stop(
    "The following required columns are missing in the Kendall tau results file:\n",
    paste(missing_cols, collapse = ", ")
  )
}

plot_df <- tau_df %>%
  dplyr::filter(
    Variable_name %in% c(
      "GDP_per_capita_2000",
      "GDP_per_capita_2019",
      "GDP_per_capita_mean_2000_2019",
      "GDP_per_capita_trend_2000_2019_pct"
    )
  ) %>%
  dplyr::mutate(
    Variable_plot = dplyr::case_when(
      Variable_name == "GDP_per_capita_2000" ~ "GDP per capita\nin 2000",
      Variable_name == "GDP_per_capita_2019" ~ "GDP per capita\nin 2019",
      Variable_name == "GDP_per_capita_mean_2000_2019" ~ "Mean GDP per capita\n2000–2019",
      Variable_name == "GDP_per_capita_trend_2000_2019_pct" ~ "Trend in GDP per capita\n2000–2019"
    ),
    Variable_plot = factor(
      Variable_plot,
      levels = c(
        "GDP per capita\nin 2000",
        "GDP per capita\nin 2019",
        "Mean GDP per capita\n2000–2019",
        "Trend in GDP per capita\n2000–2019"
      )
    ),
    tau_label = sprintf("%.3f", Kendall_tau),
    p_label = ifelse(
      is.na(P_value),
      NA_character_,
      ifelse(P_value < 0.001, "p<0.001", paste0("p=", sprintf("%.3f", P_value)))
    )
  ) %>%
  dplyr::arrange(Variable_plot)

if (nrow(plot_df) != 4) {
  stop("Expected 4 rows for plotting, but found: ", nrow(plot_df))
}

cat("\nData used for plotting:\n")
print(plot_df)

y_min <- min(0, min(plot_df$Kendall_tau, na.rm = TRUE) * 1.25)
y_max <- max(0, max(plot_df$Kendall_tau, na.rm = TRUE) * 1.25)
y_range <- y_max - y_min

plot_df <- plot_df %>%
  dplyr::mutate(
    tau_label_y = ifelse(
      Kendall_tau >= 0,
      Kendall_tau + 0.02 * y_range,
      Kendall_tau - 0.04 * y_range
    ),
    tau_vjust = ifelse(Kendall_tau >= 0, 0, 1),
    p_label_y = ifelse(
      Kendall_tau >= 0,
      0 - 0.035 * y_range,
      0 + 0.035 * y_range
    )
  )

bar_color <- "#7588F7"
p_text_color <- "black"

p <- ggplot2::ggplot(plot_df, ggplot2::aes(x = Variable_plot, y = Kendall_tau)) +
  ggplot2::geom_col(
    width = 0.62,
    fill = bar_color,
    colour = bar_color
  ) +
  ggplot2::geom_hline(yintercept = 0, linewidth = 0.5, colour = "black") +
  ggplot2::geom_text(
    ggplot2::aes(y = tau_label_y, label = tau_label, vjust = tau_vjust),
    size = 4
  ) +
  ggplot2::geom_text(
    ggplot2::aes(y = p_label_y, label = p_label),
    colour = p_text_color,
    size = 4
  ) +
  ggplot2::labs(
    x = NULL,
    y = "Kendall's tau"
  ) +
  ggplot2::coord_cartesian(ylim = c(y_min, y_max), clip = "off") +
  ggplot2::theme_bw(base_size = 13) +
  ggplot2::theme(
    panel.grid.major.x = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_text(size = 11, colour = "black", lineheight = 0.95),
    axis.text.y = ggplot2::element_text(size = 11, colour = "black"),
    axis.title.y = ggplot2::element_text(size = 13),
    axis.ticks = ggplot2::element_line(linewidth = 0.4),
    axis.ticks.length = grid::unit(-0.15, "cm"),
    plot.margin = ggplot2::margin(8, 8, 8, 8)
  )

ggplot2::ggsave(output_png, p, width = 8.5, height = 5.8, dpi = 600, bg = "white")
ggplot2::ggsave(output_pdf, p, width = 8.5, height = 5.8, bg = "white")

write.csv(plot_df, output_csv, row.names = FALSE, fileEncoding = "UTF-8")
saveRDS(plot_df, output_rds)

cat("\nBar plot with p-values created successfully.\n")
cat("PNG saved to:\n", output_png, "\n")
cat("PDF saved to:\n", output_pdf, "\n")
cat("Plot data CSV saved to:\n", output_csv, "\n")
cat("Plot data RDS saved to:\n", output_rds, "\n")
```

## 13. Kendall by country and by region

```{r}
data_dir <- dir_corr

input_main_rds <- check_file(file.path(data_dir, "Sustainable_trend_pattern_with_GDP_extracted.rds"))
input_info_rds <- check_file(file.path(data_dir, "table1_df_updated_20250307.rds"))

output_dir <- file.path(data_dir, "Analysis_9_Kendall_by_country_and_region")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

output_excel <- file.path(output_dir, "Kendall_by_country_and_region.xlsx")
output_country_csv <- file.path(output_dir, "Kendall_by_country.csv")
output_region_csv <- file.path(output_dir, "Kendall_by_region.csv")
output_merged_rds <- file.path(output_dir, "Merged_data_for_country_region_Kendall.rds")
output_results_rds <- file.path(output_dir, "Kendall_by_country_and_region.rds")

main_df <- readRDS(input_main_rds)
info_df <- readRDS(input_info_rds)

cat("Main data loaded successfully.\n")
cat("Main data columns:\n")
print(names(main_df))

cat("\nInfo data loaded successfully.\n")
cat("Info data columns:\n")
print(names(info_df))

find_first_existing_col <- function(df, candidates) {
  hit <- candidates[candidates %in% names(df)]
  if (length(hit) == 0) return(NA_character_)
  hit[1]
}

clean_text <- function(x) {
  x <- trimws(as.character(x))
  x[x %in% c("", "NA", "NaN", "NULL")] <- NA
  x
}

normalize_region_14 <- function(x) {
  x0 <- clean_text(x)
  x1 <- tolower(x0)

  out <- rep(NA_character_, length(x1))

  out[grepl("central asia", x1)] <- "Central Asia"
  out[grepl("eastern asia", x1)] <- "Eastern Asia"
  out[grepl("eastern europe", x1)] <- "Eastern Europe"
  out[grepl("latin america", x1) | grepl("caribbean", x1)] <- "Latin America and the Caribbean"
  out[grepl("northern africa", x1)] <- "Northern Africa"
  out[grepl("northern america", x1)] <- "Northern America"
  out[grepl("northern europe", x1)] <- "Northern Europe"
  out[grepl("south-eastern asia", x1) | grepl("southeastern asia", x1) | grepl("south eastern asia", x1)] <- "South-eastern Asia"
  out[grepl("southern asia", x1)] <- "Southern Asia"
  out[grepl("southern europe", x1)] <- "Southern Europe"
  out[grepl("sub-saharan africa", x1) | grepl("sub saharan africa", x1)] <- "Sub-Saharan Africa"
  out[grepl("western asia", x1)] <- "Western Asia"
  out[grepl("western europe", x1)] <- "Western Europe"

  out[grepl("oceania", x1)] <- "Oceania"
  out[grepl("australia", x1)] <- "Oceania"
  out[grepl("new zealand", x1)] <- "Oceania"
  out[grepl("other oceania", x1)] <- "Oceania"
  out[grepl("melanesia", x1)] <- "Oceania"
  out[grepl("micronesia", x1)] <- "Oceania"
  out[grepl("polynesia", x1)] <- "Oceania"

  out
}

make_na_result <- function(group_type, group_name, xvar, n_obs, status_text) {
  data.frame(
    Group_type = group_type,
    Group_name = group_name,
    Variable_name = xvar,
    N = n_obs,
    Kendall_tau = NA_real_,
    P_value = NA_real_,
    Direction = NA_character_,
    Significance = NA_character_,
    Status = status_text,
    stringsAsFactors = FALSE
  )
}

run_kendall_one_group <- function(
  data,
  xvar,
  yvar = "Pattern_code",
  group_name = NA_character_,
  group_type = NA_character_,
  min_n = 10,
  min_unique_x = 3,
  min_unique_y = 2
) {
  sub <- data[, c(xvar, yvar), drop = FALSE]
  sub <- sub[!is.na(sub[[xvar]]) & !is.na(sub[[yvar]]), , drop = FALSE]

  n_obs <- nrow(sub)

  if (n_obs < min_n) {
    return(make_na_result(group_type, group_name, xvar, n_obs, paste0("Skipped: N < ", min_n)))
  }

  ux <- length(unique(sub[[xvar]]))
  if (ux < min_unique_x) {
    return(make_na_result(group_type, group_name, xvar, n_obs, paste0("Skipped: unique x < ", min_unique_x)))
  }

  uy <- length(unique(sub[[yvar]]))
  if (uy < min_unique_y) {
    return(make_na_result(group_type, group_name, xvar, n_obs, paste0("Skipped: unique y < ", min_unique_y)))
  }

  warning_msg <- NULL

  kt <- tryCatch(
    withCallingHandlers(
      Kendall::Kendall(sub[[xvar]], sub[[yvar]]),
      warning = function(w) {
        warning_msg <<- conditionMessage(w)
        invokeRestart("muffleWarning")
      }
    ),
    error = function(e) {
      return(structure(list(.error = conditionMessage(e)), class = "kendall_error"))
    }
  )

  if (inherits(kt, "kendall_error")) {
    return(make_na_result(group_type, group_name, xvar, n_obs, paste0("Kendall error: ", kt$.error)))
  }

  tau_val <- suppressWarnings(as.numeric(kt$tau))
  p_val <- suppressWarnings(as.numeric(kt$sl))

  if (!is.finite(tau_val) || !is.finite(p_val)) {
    return(make_na_result(group_type, group_name, xvar, n_obs, "Kendall returned non-finite result"))
  }

  status_text <- if (is.null(warning_msg)) "OK" else paste0("Warning captured: ", warning_msg)

  data.frame(
    Group_type = group_type,
    Group_name = group_name,
    Variable_name = xvar,
    N = n_obs,
    Kendall_tau = tau_val,
    P_value = p_val,
    Direction = ifelse(tau_val > 0, "Positive", ifelse(tau_val < 0, "Negative", "Zero")),
    Significance = ifelse(
      p_val < 0.001, "p < 0.001",
      ifelse(p_val < 0.01, "p < 0.01", ifelse(p_val < 0.05, "p < 0.05", "Not significant"))
    ),
    Status = status_text,
    stringsAsFactors = FALSE
  )
}

main_x_col <- find_first_existing_col(main_df, c("x", "lon", "longitude", "X", "Lon", "Longitude"))
main_y_col <- find_first_existing_col(main_df, c("y", "lat", "latitude", "Y", "Lat", "Latitude"))

info_x_col <- find_first_existing_col(info_df, c("x", "lon", "longitude", "X", "Lon", "Longitude"))
info_y_col <- find_first_existing_col(info_df, c("y", "lat", "latitude", "Y", "Lat", "Latitude"))

pattern_col <- find_first_existing_col(main_df, c("name", "pattern", "Pattern", "Sustainable_trend_pattern", "trend_pattern"))
pattern_code_col <- find_first_existing_col(main_df, c("Pattern_code", "pattern_code", "code", "Code"))

country_col <- find_first_existing_col(
  info_df,
  c("Country", "country", "country_name", "Country_name", "NAME_0", "admin0", "nation", "Nation")
)

region_col <- "SUBREGION2"

cat("\nDetected columns:\n")
cat("main_x_col =", main_x_col, "\n")
cat("main_y_col =", main_y_col, "\n")
cat("info_x_col =", info_x_col, "\n")
cat("info_y_col =", info_y_col, "\n")
cat("pattern_col =", pattern_col, "\n")
cat("pattern_code_col =", pattern_code_col, "\n")
cat("country_col =", country_col, "\n")
cat("region_col =", region_col, "\n")

if (is.na(main_x_col) || is.na(main_y_col)) stop("Could not detect x/y columns in main_df.")
if (is.na(info_x_col) || is.na(info_y_col)) stop("Could not detect x/y columns in info_df.")
if (is.na(pattern_col)) stop("Could not detect Sustainable trend pattern column in main_df.")
if (is.na(country_col)) stop("Could not detect country column in info_df.")
if (!(region_col %in% names(info_df))) stop("The fixed region column SUBREGION2 does not exist in info_df.")

analysis_vars <- c(
  "GDP_per_capita_2000",
  "GDP_per_capita_2019",
  "GDP_per_capita_mean_2000_2019",
  "GDP_per_capita_trend_2000_2019_pct"
)

missing_gdp_cols <- analysis_vars[!analysis_vars %in% names(main_df)]
if (length(missing_gdp_cols) > 0) {
  stop(
    "The following GDP columns are missing in main_df:\n",
    paste(missing_gdp_cols, collapse = ", ")
  )
}

main_df[[main_x_col]] <- as.numeric(main_df[[main_x_col]])
main_df[[main_y_col]] <- as.numeric(main_df[[main_y_col]])

main_df[[pattern_col]] <- clean_text(main_df[[pattern_col]])
main_df[[pattern_col]] <- toupper(main_df[[pattern_col]])

for (v in analysis_vars) {
  main_df[[v]] <- as.numeric(main_df[[v]])
}

if (!is.na(pattern_code_col) && pattern_code_col %in% names(main_df)) {
  main_df$Pattern_code <- as.integer(main_df[[pattern_code_col]])
} else {
  main_df$Pattern_code <- ifelse(
    main_df[[pattern_col]] == "SD", 1,
    ifelse(
      main_df[[pattern_col]] == "MD", 2,
      ifelse(
        main_df[[pattern_col]] == "MT", 3,
        ifelse(main_df[[pattern_col]] == "MI", 4, ifelse(main_df[[pattern_col]] == "SI", 5, NA))
      )
    )
  )
}

main_df <- main_df[!is.na(main_df$Pattern_code) & main_df$Pattern_code %in% 1:5, , drop = FALSE]

main_df$x_merge <- round(main_df[[main_x_col]], 6)
main_df$y_merge <- round(main_df[[main_y_col]], 6)

info_df[[info_x_col]] <- as.numeric(info_df[[info_x_col]])
info_df[[info_y_col]] <- as.numeric(info_df[[info_y_col]])

info_df$Country_name <- clean_text(info_df[[country_col]])
info_df$Region_raw <- clean_text(info_df[[region_col]])
info_df$Region_14 <- normalize_region_14(info_df$Region_raw)

info_df$x_merge <- round(info_df[[info_x_col]], 6)
info_df$y_merge <- round(info_df[[info_y_col]], 6)

info_keep <- info_df[, c("x_merge", "y_merge", "Country_name", "Region_raw", "Region_14"), drop = FALSE]
info_keep <- unique(info_keep)

merged_df <- merge(main_df, info_keep, by = c("x_merge", "y_merge"), all.x = TRUE)

cat("\nMerged data rows:", nrow(merged_df), "\n")
cat("Rows with country info:", sum(!is.na(merged_df$Country_name)), "\n")
cat("Rows with 14-region info:", sum(!is.na(merged_df$Region_14)), "\n")

saveRDS(merged_df, output_merged_rds)

country_df <- merged_df[!is.na(merged_df$Country_name), , drop = FALSE]
country_names <- sort(unique(country_df$Country_name))

country_results_list <- list()
counter <- 1

for (ct in country_names) {
  sub_ct <- country_df[country_df$Country_name == ct, , drop = FALSE]

  for (v in analysis_vars) {
    country_results_list[[counter]] <- run_kendall_one_group(
      data = sub_ct,
      xvar = v,
      yvar = "Pattern_code",
      group_name = ct,
      group_type = "Country",
      min_n = 10,
      min_unique_x = 3,
      min_unique_y = 2
    )
    counter <- counter + 1
  }
}

country_results <- do.call(rbind, country_results_list)

region_df <- merged_df[!is.na(merged_df$Region_14), , drop = FALSE]
region_names <- c(
  "Central Asia",
  "Eastern Asia",
  "Eastern Europe",
  "Latin America and the Caribbean",
  "Northern Africa",
  "Northern America",
  "Northern Europe",
  "Oceania",
  "South-eastern Asia",
  "Southern Asia",
  "Southern Europe",
  "Sub-Saharan Africa",
  "Western Asia",
  "Western Europe"
)

region_results_list <- list()
counter <- 1

for (rg in region_names) {
  sub_rg <- region_df[region_df$Region_14 == rg, , drop = FALSE]

  for (v in analysis_vars) {
    region_results_list[[counter]] <- run_kendall_one_group(
      data = sub_rg,
      xvar = v,
      yvar = "Pattern_code",
      group_name = rg,
      group_type = "Region",
      min_n = 10,
      min_unique_x = 3,
      min_unique_y = 2
    )
    counter <- counter + 1
  }
}

region_results <- do.call(rbind, region_results_list)

var_label_map <- c(
  "GDP_per_capita_2000" = "GDP per capita in 2000",
  "GDP_per_capita_2019" = "GDP per capita in 2019",
  "GDP_per_capita_mean_2000_2019" = "Mean GDP per capita, 2000-2019",
  "GDP_per_capita_trend_2000_2019_pct" = "Trend in GDP per capita, 2000-2019 (%/year)"
)

country_results$Variable <- unname(var_label_map[country_results$Variable_name])
region_results$Variable <- unname(var_label_map[region_results$Variable_name])

country_results <- country_results[, c(
  "Group_type", "Group_name", "Variable", "Variable_name",
  "N", "Kendall_tau", "P_value", "Direction", "Significance", "Status"
), drop = FALSE]

region_results <- region_results[, c(
  "Group_type", "Group_name", "Variable", "Variable_name",
  "N", "Kendall_tau", "P_value", "Direction", "Significance", "Status"
), drop = FALSE]

wb <- openxlsx::createWorkbook()

openxlsx::addWorksheet(wb, "Country_Kendall")
openxlsx::writeData(wb, "Country_Kendall", country_results)

openxlsx::addWorksheet(wb, "Region_Kendall")
openxlsx::writeData(wb, "Region_Kendall", region_results)

openxlsx::addWorksheet(wb, "README")
openxlsx::writeData(
  wb, "README",
  data.frame(
    Item = c(
      "Main data file",
      "Info data file",
      "Pattern coding",
      "Country grouping",
      "Region grouping",
      "Fixed region column",
      "Oceania handling",
      "Method",
      "Merge method",
      "Robust filtering",
      "Status column"
    ),
    Description = c(
      basename(input_main_rds),
      basename(input_info_rds),
      "SD=1, MD=2, MT=3, MI=4, SI=5",
      "Kendall run separately for each country",
      "Kendall run separately for 14 target regions",
      "SUBREGION2",
      "Australia, New Zealand, other Oceania, and related labels are aggregated into Oceania",
      "Kendall::Kendall with warning/error capture",
      "Merged by rounded coordinates (6 decimals)",
      "Groups with N < 10, too few unique x, or too few unique y are skipped and returned as NA",
      "Status records whether result is OK, skipped, or warning/error captured"
    ),
    stringsAsFactors = FALSE
  )
)

openxlsx::saveWorkbook(wb, output_excel, overwrite = TRUE)

write.csv(country_results, output_country_csv, row.names = FALSE, fileEncoding = "UTF-8")
write.csv(region_results, output_region_csv, row.names = FALSE, fileEncoding = "UTF-8")

saveRDS(
  list(
    merged_data = merged_df,
    country_results = country_results,
    region_results = region_results
  ),
  output_results_rds
)

cat("\nAnalysis 9 completed successfully.\n")
cat("Merged data RDS saved to:\n", output_merged_rds, "\n")
cat("Excel saved to:\n", output_excel, "\n")
cat("Country CSV saved to:\n", output_country_csv, "\n")
cat("Region CSV saved to:\n", output_region_csv, "\n")
cat("Results RDS saved to:\n", output_results_rds, "\n")
```